\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
%----------------------------------------------------------------------------------------

Choosing this topic for my master thesis has been a leap in the dark: I knew nothing about microRNAs and very little about biology in general. However, curiosity and the feeling that, for once, I could do something useful, made this decision pretty straightforward. I have to admit, though, that it has been more difficult and stressful than I thought.

Data was inconsistent and mostly unstructured, papers and articles about the subject confusing and too technical for someone who hardly knew what a cell was. Nevertheless, everybody kept encouraging and supporting me during this trip and I would like to thank them all for this.

Foremost I thank my advisor Sebastiano Vigna (Seba) for his unusual availability and patience in trying to help me out with this topic, even from the other side of the world. I really appreciated his constant support and his unwavering trust in me.  

A big thanks goes to the whole LAW laboratory and especially to Giorgio, who now lives in London, for all lunches and breaks together.

I also thank my parents and my family for the unceasing encouragement and for supporting me during this venture.

Thanks to my best friends Gianluca, Diego and Stefano for always reminding me that there is more to life than just work and study and for the uncountable dinners, birthday's parties and other outings together.

Finally I want to thank all my university colleagues: Daniele, Massimiliano, Chiara and everyone else who had the patience to stand me during this 5 years: who would have thought, 10 years ago, that in 2019 I'd still be in university!! 